\LARGE \bf How Children Perceive and Interact with a Robot that Behaves Unexpectedly -- The Domino Experiment

### 2014-09-15 paper.tex
First pass on the whole article. 11 pages for now

\title{\LARGE \bf
How Children Perceive and Interact with a Robot that Behaves Unexpectedly - The Domino Experiment
}

@@

\textbf{Engagement} is a metric that has been extensively used and studied both
in HRI and interactions with other agent-like systems. It has been defined from
several perspectives. For example \cite{sidner_where_2004} define engagement as
\textit{``the process by which two (or more) participants establish, maintain
and end their perceived connections''}. A definition of long-term engagement is
proposed by \cite{bickmore_maintaining_2010}: \textit{``the degree of
involvement a user chooses to have with a system over time''}.

Different possibilities to foster engagement (both short- and long-term
engagement) in HRI have been explored, in particular with social robots. A lot
of research has moved toward creating sophisticated emotional models which cause
complex robot behavior. Some other work
\cite{bickmore_maintaining_2010,short_no_2010} has shown that there can be much
simpler ways to enhance engagement.

\cite{bickmore_maintaining_2010} describe a series of longitudinal studies on
engagement with an agent-like system. They demonstrated that user engagement
with an interface agent can be increased using relatively simple techniques and
manipulations that make the agent more life-like and human. For instance, when
the agent showed variations in its behavior, participants were more engaged and
reported a desire to continue interacting with the agent.

Similarly, more looking at short-term engagement, \cite{short_no_2010} found
that a simple manipulation of the robot's behavior can lead to greater
engagement. The authors let participants play several rounds of the
rock-paper-scissors game with the robot (the playfulness of the scenario seems
important). When the robot was cheating from time to time, participants tended
to ascribe intention to the robot what in turn led to greater engagement in how
they were interacting with the robot. The mechanism behind the cheating action
that the robot showed, was that participants did not expect this behavior from
the robot. The authors suggest that \textit{``any deviation from expected
operation is sufficient to create a greater degree of engagement in the
interaction.''} Moreover, \cite[p.~225]{short_no_2010} proposed that
\textit{``many interactions can be improved by the introduction of such simple
behaviors, and that this should be exploited by designers in HRI. Bringing human
and robot together to perform a simple, repetitive, familiar task and then
having the robot behave unexpectedly can increase engagement and mental state
attribution without complex behavioral or mechanical additions.''}
\cite{leite_long-term_2013} studied the long-term engagement of children with a
chess playing robot that adapted its behavior to the children and showed empathy
toward them. She found that empathetic robots are more likely to engage users in
the long-term and they proposed several guidelines for designing such artificial
companions. Our aim for the ``Domino Study'' is however less ambitious than
creating a long-term robot companion. We are still at an early stage of the
development and are using a remote controlled prototype as we did in the
previous study. What we would like to achieve in this study is to find out how
we can enhance children's experience with the robot so that they do not find it
repetitive and boring to interact with Ranger. In general, repetitiveness is
likely to decrease a user's motivation to continue using a system
\cite{bickmore_establishing_2005}.

We present in this article a study that explores possibilities of sustaining
children's engagement with the Ranger~\cite{mondada2014ranger} robot by
manipulating the robot's behavior in such way that it appears
\textit{unexpected} to the children. We examine how different variations of
robot behavior impact children's interaction with Ranger and their perception of
it (\eg in terms of attributing intention and cognitive abilities to the robot).

@@

@@

    \item The robot makes a \textbf{mistake}: When called by the child to come
    over, the robot goes wrong but recognizes its mistake and repairs. We expect
    this to be perceived (explicitly) as \textit{``to err is human''}, and
    (implicitly) as the robot being endowed with a certain level of
    introspective capabilities (it was able to recognize its own error). In this
    condition, we assume increased attributions of human-likeness to the
    robot.\footnote{The attribution of human-likeness to a robot is labeled as
    \textit{anthropomorphism}, and we will hereafter sometimes use ``attribute
    intention'' and ``anthropomorphize'' synonymously.}

@@

We analyzed children's reaction focusing on two main aspects. On one hand,
children's \textbf{behaviour} (their reactions) toward the unexpected robot
behaviour was studied in terms of \textbf{active engagement} with the robot. On
the other hand, we analyzed children's \textbf{perception} of the robot in term
of \textbf{anthropomorphism} -- the attribution of human-like characteristics,
such as cognitive abilities and the ability to show intentions. We assumed that
in general a robot that behaves unexpectedly from time to time can promote
engagement and lead children to attribute intention to it. \cite{short_no_2010}
found that participants anthropomorphize a cheating robot more than a robot that
always behaves fairly, and also evaluate the interaction as more engaging. Based
on the related work we formulate the following two hypotheses:

@@

    \item[Hypothesis 2:] Children perceive a robot that shows intention or
        cognitive abilities as more human-like than a robot that appears to have
        a system error, \ie the disobeying robot and the robot that makes a
        mistake will be more anthropomorphized than the robot that gets lost
        (between subjects variable).

@@

Our research questions deal with both children's observable behaviour and their
perception of the robot. We would also like to explore the relation between
these two aspects, related to the attribution of human-like characteristics to
the robot. The motivation behind this is to find out whether
\textit{anthropomorphism} can not only be measured as a specific type of
perception but also as an observable behaviour in the interaction itself. Thus,
we would like to bridge things together and, relying on the concept of an
\textbf{anthropomorphism index} introduced by~\cite{fink2014dynamics}, build a
compound metric that considers both children's perception and interaction
aspects (qualitatively). Previous work suggests that a social relation to a
robot (we view anthropomorphism as a specific type of social relation) reflects
an increased engagement and can be effective in sustaining interaction.
Consequently, we formulate a third hypotesis:

@@

We used the same prototype as in the previous study, and Ranger was again
controlled by a human Wizard, who was in the same room (see
Figure~\ref{fig:domino-setup}). This was to ensure that the Wizard could see
children's interaction and hear their commands to the robot without any delay.

@@

The basic idea of the interaction scenario was that there are two children who
play domino together, with the help of the robotic box Ranger. The scenario
setup is displayed in Figure~\ref{fig:domino-setup}. The challenge is that the
tiles of the domino are distributed in the room, hidden behind three beanbags,
and while one child -- \textit{the searcher} -- searches for the right tile, the
other child -- \textit{the receiver} -- is asked to stay in a play tent. There
is a ``river'' (play carpet) in between the two children that we told them they
cannot cross and therefore they need the robot to transport the domino tiles
between them.

@@

To start the game (divided in several \textit{runs} that correspond each to the
delivery and assembling of one domino tile), there is already one domino tile in
front of the tent, where \textit{the receiver} child stays and assembles the
domino chain. \textit{The receiver} child asks \textit{the searcher} child for a
specific tile, \eg a tile with a donkey, \textit{the searcher} searches for the
respective tile, sits down on the next beanbag, and asks the robot to come over.
The Ranger robot is first located next to the tent, then, when called by the
\textit{searcher} it starts moving, crosses the river carpet, and comes over to
the \textit{searcher} on the beanbag. The \textit{searcher} child puts the
domino tile into the robotic box, and the robot then goes back to \textit{the
receiver} child in the tent. Then, \textit{the receiver} takes out the domino
tile from the robot, and puts the two tiles together. The first \textit{run} is
over, and a new \textit{run} starts, when \textit{the receiver} asks \textit{the
searcher} for the next domino tile.

@@

In our study, there were a total of 14 runs, and at each run, the robot exhibits
one out of four possible behaviours: \emph{correct}, \emph{mistake}, \emph{lost}
or \emph{disobey}.  The first 5 runs were used to set the baseline and the robot
always behaved correctly. In the 9 remaining runs, the robot showed one of the
misbehaviour (\emph{mistake}, \emph{lost} or \emph{disobey}) at the $3^{rd}$ and
$4^{th}$ run as well as at the $7^{th}$ and $8^{th}$ run. The study is build as
a between-subject study, and the type of misbehaviour was therefore always the
same for a given group.

Special attention has been paid to the distribution of misbehaving runs during
the study. The five first correct runs aims at setting the children expectation
regarding a consistent robot behaviour. When the robot then exhibits an
unexpected behaviour, children are likely to be positively surprised. A similar
effect has been observed in a study with a robot that was cheating from time to
time~\cite{short_no_2010}.  Besides, to prevent the robot misbehaviours to be
immediately interpreted as failures, we introduce these misbehaviour neither at
the very beginning nor at very end of the interaction
(\cite{desai_effects_2012,desai_impact_2013} show that early and late robot
failures negatively impact trust).

\cite{short_no_2010} also adopted a similar pattern when introducing unexpected
robot behaviour. In their study with a robot that cheated, from the 20 rounds of
the rock-paper-scissors game, the robot cheated three times in the middle: on
the 4th, 8th and 15th round. The authors adjusted the interaction length so that
the spacing and timing of the cheating was preserved.

@@

In the \emph{correct} robot behaviour (baseline behaviour) consists in the
following states:

@@

@@

Overall, there were 13 pairs of children (n=26) participating in the interaction
study: 16 boys and 10 girls, 4-5 years old (M=4.46, SD=0.45). In 11 of the
pairs, children were friends who knew each other from kindergarten, nursery
school or because they lived in the same neighborhood. 2 of the pairs were
composed of brother and sister.

@@

We have obtained and analysed two types of data. The \textbf{children's
behaviours toward the robot} (\ie the child-robot interactions) have been
captured in the video recordings by annotating a set of actions, described
below. The children's \textbf{perception of the
robot} has been captured through audio-recorded semi-structured interviews.

@@

We annotated children's behavior in the video records, and coded 10 different
actions:

@@

In total, we obtained 2354 distinct actions which summed up to a total
annotation duration of 145 minutes. On average, one child accounted for 92
actions (SD=23). Figure~\ref{fig:domino-time-all} shows the distribution of
actions over the different runs, in any condition.

\begin{figure*}[ht!] 
    \centering 
    \includegraphics[width=0.8\linewidth]{domino-time-all.pdf} 
    \caption{\small \textbf{Number and type of actions for each run} (average for
    one child per minute). Generally, the number of actions does not decrease over
    time (from run 1.1 to run 2.9). The first 7 runs correspond to the
    \textit{expected phase}, the second 7 runs correspond to the \textit{unexpected
    phase}. Especially during run 2.3, the first time when the robot showed an
    unexpected behavior, children tended to \textit{look} more at the experimenter.
    During the unexpected phase, also \textit{talk} and \textit{gesture} seem to be
    increased.}

    \label{fig:domino-time-all} 
\end{figure*}

The actions \textit{put, remove} and \textit{call} were ``requested'' actions
because the scenario required them and children were asked to carry them out.
Hence, these actions are not relevant to be analyzed quantitatively. The other
actions \textit{explore, misuse, gesture, touch, show, talk} and \textit{look}
are spontaneous actions that were not requested but arose in the interaction.
These actions reflect engagement, and are the only one we considered during the
analysis (except otherwise indicated).

@@

It is not easy to interview young children: they have a short attention span, we
(the experimenters) are strangers to them, and they are in an unfamiliar
environment. The pre-interview served to ``break the ice'' and during the first
playful interaction phase, most children familiarized themselves with the
situation.

Also, children do not articulate themselves and understand questions like
adults. We tested our interview structure and the script in a pilot-study with
three young children, and were assisted in formulating our questions by a
pedagogue and father of four. This helped us adapting our language and
question-style to the 4-5 year old participants. We set up the interviews like a
casual conversation / discussion, so we did not separate the two children in
order to keep the situation natural. We used casual and informal language
adapted to the age of the children, and encouraged them to justify their answers
and tell us more details by asking, for example, \textit{How do you know ... ?}
or \textit{Tell me, what have you observed when ... ?}. We paid attention to not
``put words in children's mouth''. Consequently, though we re-phrased and
repeated some questions, we never forced children to give an answer, and
accepted when they said they would not know or when they did not respond at all.

\begin{table}[h]
\centering
\footnotesize
\begin{tabular}{p{0.9\linewidth}}
    \toprule
    \textbf{Construct \emph{expectation}} \\
    \midrule

@@

    %%
    \toprule
    \textbf{Ascribe cognitive abilities} \\
    \midrule

@@

    \caption{\small Questions used during the semi-structured interviews
    with children. The Ranger robot toy box is abbreviated with ``R''.}

@@

We generally needed to keep the interviews as short as possible. In designing
our interview script and selecting relevant questions, we took inspiration from
previous work on child-robot interaction and children's perception of robots
\cite{kahn_jr._robotic_2006,weiss_i_2009,leite_influence_2013}. For instance, we
applied and adapted some of the ``constructs'' and example questions from the
questionnaires used in \cite{kahn_jr._robotic_2006} and \cite{weiss_i_2009}. The
authors evaluated children's perception of the robotic dog AIBO after the
children had played with the robot. A ``construct'' addresses a specific factor
(topic) that can be measured by several questions. For instance, the construct
``cognitive abilities'' (called ``cognition'' in \cite{weiss_i_2009}) considers
the robot's ability to hear and to see (perceptual skills), as attributed by the
children. The construct ``moral standing'' and the respective question was taken
from \cite{kahn_jr._robotic_2006}.\footnote{According to
\cite{kahn_jr._robotic_2006}, \textit{moral} refers to considerations based on
an artifact's physical or psychological welfare, and virtue (whether the
artifact deserves care). An attribution of moral standing reflects, for
instance, that the robot engenders moral regard, is morally responsible,
blameworthy, has rights or deserves respect.}			Similarly, we grouped
questions according to specific constructs that they evaluate (see
Table~\ref{tab:domino-questions}). This was an adaption and extension of the
questions and constructs used in previous work.\\

@@

Similar to how we processed interviews in the previous studies, we qualitatively
transcribed interviews, \ie we did not craft a full word-by-word transcript but
noted down any statement that was useful and relevant. This was organized in a
spreadsheet, so to obtain an overview of the different replies to a question.
The interviews were analyzed in a qualitative manner.

@@

we can investigate how far these two types of data can
be used to understand children's engagement in the interaction and how far they
anthropomorphize the robot. We would like to develop a toolkit that considers
both children's perception and interaction, in order to measure anthropomorphism
(as a special type of human-like engagement with the robot).

There are different possibilities to measure engagement in HRI, depending on the
specific research question and context. Metrics to measure behavioral engagement
in the interaction include, for instance, conversation analysis (\eg used in
\cite{short_no_2010}) or general attention analysis. These can be studied by
analyzing interaction videos in terms of head movement, eye tracking and gesture
/ body movement analysis (\eg in \cite{sidner_explorations_2005}).
Post-measurements to measure emotional engagement can be questionnaires that try
to assess constructs like the perceived presence and involvement in the
interaction. An example is the \textit{Interactive Experience Questionnaire}
(originally developed by \cite{lombard_measuring_2000}) of which adapted
versions were used in
\cite{kidd_effect_2004,bainbridge_effect_2008,short_no_2010}.

A mix of several methods has been used in a long-term interaction study with 8-9
year old children, \cite{leite_long-term_2013}. The author measured engagement
through video observations (by analyzing the amount of time that children spent
looking at the robot), interviews, and questionnaires. The interviews were
semi-structured, containing initial yes-or-no questions followed by open-ended
questions that allowed children to justify and elaborate their answers. We have
chosen a similar approach that considers both children's behavioral and
emotional engagement.

With 4-5 year old children, however, we cannot use rating scales to ask them
about constructs as abstract as social presence or how much they felt involved
in the interaction. Therefore we have attempted to combine \emph{interaction}
data with the results of the semi-structured interviews.  Several of the
aforementioned coded actions can reflect engagement: \textit{explore, misuse,
gesture, touch, show,} and \textit{talk}. Also, \textit{look} at the
experimenter can be considered as engagement in the interaction: We assume that
children look at the experimenter when they are surprised and seek for help.
This behavior reflects that they want to make sense of the robot's behavior and
that they notice it as ``strange''. Further, we can take into account how they
refer to Ranger and describe the robot (as well as their experience) in the
interviews.  Specifically, we used the constructs \textit{ascribe intention,
    ascribe cognitive abilities, ascribe emotional state}, and \textit{ascribe
    moral standing} (Table~\ref{tab:domino-questions}) to assess how far they
    anthropomorphize the robot.

To account for the fact that anthropomorphism arises in an interaction, we try
to bring both children's perception of the robot (post-measurement) and their
behavior toward it (in-the-moment measurement) together. By doing so, we would
like to obtain a qualitative \textbf{anthropomorphism
index}~\cite{fink2014dynamics}. We quantify the index by giving 1 point for each
anthropomorphic perception of the robot (max.  13 points) and for specific kinds
of human-like behavior toward the robot (max.  3 points). The index included the
following aspects:\\

\textbf{Perception: \textit{(max. 13 points)}}
\begin{itemize}
	\item Ascribe \textbf{mental states / feelings} to Ranger: \textit{0-4 points}\\
	(2 points for agreeing that Ranger can be happy or sad; 2 points for attributing Ranger with hunger or tiredness)
	\item Ascribe \textbf{cognitive abilities / intention} to Ranger: \textit{0-4 points}\\ 
	(each 0.5 points for ascribing seeing and hearing ability; 1 point for agreeing that Ranger can go out the door by itself; 1 point for disagreeing that Ranger always obeys; 1 point for agreeing that Ranger can do something silly)
	\item Ascribe \textbf{sociality / companionship} to Ranger: \textit{1 point}\\ 
	(1 point for agreeing that Ranger can be a friend)
	\item Ascribe \textbf{moral standing} to Ranger: \textit{1 point}\\ (1 point for disagreeing that Ranger be left alone at home)
	\item Other \textbf{anthropomorphic statements}: \textit{0-3 points}\\ 
	(1 point for anthropomorphic reason for Ranger's misbehavior; 2 points for anthropomorphic reason for not leaving Ranger alone) 
\end{itemize}

\textbf{Behavior: \textit{(max. 3 points)}}
\begin{itemize}
	\item Use of \textbf{direct speech} toward Ranger: \textit{1 point}\\
	(\textit{calling} the robot to come over is not considered)
	\item Use of \textbf{polite formulations} toward Ranger: \textit{1 point}\\ 
	(\eg saying \textit{``thank you Ranger''} or \textit{``please Ranger ...''} or \textit{``goodbye''})
	\item Use of \textbf{social or pointing gestures} toward Ranger: \textit{1 point}\\ 
	(\eg waving at the robot, nodding)
\end{itemize}

>>>>

\title{\LARGE \bf
How Children Perceive and Interact with a Robot that Behaves Unexpectedly -- The Domino Experiment
}

@@

\textbf{Engagement} is a metric that has been extensively used and studied both
in HRI and during interactions with other agent-like systems. It has been
defined from several perspectives. For example \cite{sidner_where_2004} define
engagement as \textit{``the process by which two (or more) participants
establish, maintain and end their perceived connections''}. A definition of
long-term engagement is proposed by \cite{bickmore_maintaining_2010}:
\textit{``the degree of involvement a user chooses to have with a system over
time''}.

Different possibilities to foster engagement (both short- and long-term
engagement) in HRI have been explored, in particular with social robots. A lot
of research has moved toward creating sophisticated emotional models which cause
complex robot behaviour. \cite{leite_long-term_2013} studied the long-term
engagement of children with a chess playing robot that adapted its behaviour to
the children and showed empathy toward them. She found that empathetic robots
are more likely to engage users in the long-term and they proposed several
guidelines for designing such artificial companions. Some other work
\cite{bickmore_maintaining_2010,short_no_2010} has shown that there can be much
simpler ways to enhance engagement.

\cite{bickmore_maintaining_2010} describe a series of longitudinal studies on
engagement with an agent-like system. They demonstrated that user engagement
with an interface agent can be increased using relatively simple techniques and
manipulations that make the agent more life-like and human. For instance, when
the agent showed variations in its behaviour, participants were more engaged and
reported a desire to continue interacting with the agent.

Similarly, more looking at short-term engagement, \cite{short_no_2010} found
that a simple manipulation of the robot's behaviour can lead to greater
engagement. The authors let participants play several rounds of the
rock-paper-scissors game with the robot (the playfulness of the scenario seems
important). When the robot was cheating from time to time, participants tended
to ascribe intention to the robot what in turn led to greater engagement in how
they were interacting with the robot. The mechanism behind the cheating action
that the robot showed, was that participants did not expect this behaviour from
the robot. The authors suggest that \textit{``any deviation from expected
    operation is sufficient to create a greater degree of engagement in the
interaction.''} Moreover, \cite[p.~225]{short_no_2010} proposed that
\textit{``many interactions can be improved by the introduction of such simple
    behaviours, and that this should be exploited by designers in HRI. Bringing
    human and robot together to perform a simple, repetitive, familiar task and
then having the robot behave unexpectedly can increase engagement and mental
state attribution without complex behavioural or mechanical additions.''}.
In their model of the dynamics of anthropomorphism, Fink et
al.~\cite{fink2014dynamics} also suggest that \emph{disruptive behaviours}
may lead to increased anthropomorphic projections and possibly engagement.

Based on these previous researches, we would like to find out in this study how
we can enhance children's experience with the robot so that they do not find it
repetitive and boring, and ultimately keep interacting with the robot (as
repetitiveness is likely to decrease a user's motivation to continue using a
system~\cite{bickmore_establishing_2005}).

We present in this article a study that explores possibilities of sustaining
children's engagement with the \emph{Ranger}~\cite{mondada2014ranger} robot by
manipulating the robot's behaviour in such way that it appears
\textit{unexpected} to the children. We examine how different variations of
robot behaviour impact children's interaction with Ranger and their perception of
it (\eg in terms of attributing intention and cognitive abilities to the robot).

@@

This study also aims at applying to an actual study and assessing the usefulness
of the \emph{anthropomorphic index} proposed by Fink et
al.~\cite{fink2014dynamics} as a compound (behavioural and phenomenological)
metric for the attribution of human-like features to robot. We are especially
interested in investigating how this indicator predicts and/or reflects actual
\emph{engagement} into the interaction.

@@

    \item The robot makes a \textbf{mistake}: When called by the child to come
        over, the robot goes wrong but recognizes its mistake and repairs. We
        expect this to be perceived (explicitly) as \textit{``to err is
        human''}, and (implicitly) as the robot being endowed with a certain
        level of introspective capabilities (it was able to recognize its own
        error). In this condition, we assume increased attributions of
        human-likeness to the robot.\footnote{The attribution of human-likeness
            to a robot is hereafter labeled as \textit{anthropomorphic
        projections}.}

@@

We analyzed children's reaction focusing on two main aspects. On one hand,
children's \textbf{behaviour} (their reactions) toward the unexpected robot
behaviour was studied in terms of \textbf{active engagement} with the robot. On
the other hand, we analyzed children's \textbf{perception} of the robot in term
of \textbf{anthropomorphism} -- the attribution of human-like characteristics,
such as cognitive abilities and the ability to show intentions. We assumed that
in general a robot that behaves unexpectedly from time to time can promote
engagement and lead children to attribute intention to it. Based
on the related work we formulate the following two hypotheses:

@@

    \item[Hypothesis 2:] Children perceive a robot that (tentatively) displays
        intention or cognitive abilities as more human-like than a robot that
        appears to have a system error, \ie the disobeying robot and the robot
        that makes a mistake will be more anthropomorphized than the robot that
        gets lost (between subjects variable).

@@

Our research questions deal with both children's observable behaviour and their
perception of the robot. We would also like to explore the relation between
these two aspects, related to the attribution of human-like characteristics to
the robot. The motivation behind this is to find out whether
\textit{anthropomorphism} can not only be measured as a specific type of
perception but also as an observable behaviour in the interaction itself. Thus,
we would like to bridge things together and, relying on the concept of an
\textbf{anthropomorphism index}, build a
compound metric that considers both children's perception and interaction
aspects (qualitatively). Previous work suggests that a social relation to a
robot (we view anthropomorphism as a specific type of social relation) reflects
an increased engagement and can be effective in sustaining interaction.
Consequently, we formulate a third hypotesis:

@@

@@

The interaction scenario consisted in two children who play domino together,
with the help of the robotic box Ranger. The scenario setup is displayed in
Figure~\ref{fig:domino-setup}. The challenge is that the tiles of the domino are
distributed in the room, hidden behind three beanbags, and while one child --
\textit{the searcher} -- searches for the right tile, the other child --
\textit{the receiver} -- is asked to stay in a play tent. There is a ``river''
(play carpet) in between the two children that we told them they cannot cross
and therefore they need the robot to transport the domino tiles between them.

The robot was controlled by a human wizard, who was in the same room (see
Figure~\ref{fig:domino-setup}). This was to ensure that the wizard could see
children's interaction and hear their commands to the robot without any delay.
Only one group asked at the beginning of the experiment if the wizard was the
one actually controlling the robot. This did not seem to subsequently influence
their behaviour.

@@

To start the game (divided in several \emph{runs} that correspond each to the
delivery and assembling of one domino tile), there is already one domino tile in
front of the tent, where the receiver child stays and assembles the domino
chain. The receiver child asks the searcher child for a specific tile, \eg a
tile with a donkey, the searcher searches for the respective tile, sits down on
the next beanbag, and asks the robot to come over.  The Ranger robot is first
located next to the tent, then, when called by the searcher it starts moving,
crosses the river carpet, and comes over to the searcher on the beanbag. The
searcher child puts the domino tile into the robotic box, and the robot then
goes back to the receiver child in the tent. Then, the receiver takes out the
domino tile from the robot, and puts the two tiles together. The first
\emph{run} is over, and a new \emph{run} starts, when the receiver asks the
searcher for the next domino tile.

@@

In our study, there were a total of 14 runs, and at each run, the robot exhibits
one out of four possible behaviours: \emph{correct}, \emph{mistake}, \emph{lost}
or \emph{disobey}.  The first 5 runs (\emph{1.1} to \emph{1.5}) were used to set
the baseline and the robot always behaved correctly. The children then switched
the roles and in the 9 remaining runs (\emph{2.1} to \emph{2.9}), the robot
showed one of the misbehaviour (\emph{mistake}, \emph{lost} or \emph{disobey})
at the $3^{rd}$ and $4^{th}$ run as well as at the $7^{th}$ and $8^{th}$ run
(see Figure~\ref{fig:domino-time-active}). The study is build as a
between-subject study, and the type of misbehaviour was therefore always the
same for a given group.

Special attention has been paid to the distribution of misbehaving runs during
the study. The five first correct runs aims at setting the children expectation
regarding a consistent robot behaviour. When the robot then exhibits an
unexpected behaviour, children are likely to be positively surprised. A similar
effect has been observed in a study with a robot that was cheating from time to
time~\cite{short_no_2010}.  Besides, to prevent the robot misbehaviours to be
immediately interpreted as failures, we introduce these misbehaviour neither at
the very beginning nor at very end of the interaction
(\cite{desai_effects_2012,desai_impact_2013} show that early and late robot
failures negatively impact trust). \cite{short_no_2010} also adopted a similar
pattern when introducing unexpected robot behaviour. In their study with a robot
that cheated, from the 20 rounds of the rock-paper-scissors game, the robot
cheated three times in the middle: on the 4th, 8th and 15th round. The authors
adjusted the interaction length so that the spacing and timing of the cheating
was preserved.

@@

The \emph{correct} robot behaviour (baseline behaviour) consists in the
following states:

@@

In the human-repaired conditions, the experimenter instructs the searcher child
(after ~10 sec) to bring him/herself the domino tile to the robot.

@@

Overall, there were 13 pairs of children (n=26) participating in the interaction
study: 16 boys and 10 girls, 4-5 years old (M=4.46, SD=0.45), all
French-speaking. In 11 of the pairs, children were friends who knew each other
from kindergarten, nursery school or because they lived in the same
neighborhood. 2 of the pairs were composed of brother and sister.

@@

We have obtained and analysed two types of data. The \textbf{children's
behaviours toward the robot} (\ie the child-robot interactions) have been
captured in the video recordings by annotating a set of actions, described
below. The children's \textbf{phenomenological perception of the robot} has been
captured through audio-recorded semi-structured interviews which took place
between run \emph{1.5} and \emph{2.1} and at the end of the 14 runs.

@@

We annotated children's behaviour in the video records, and coded 10 different
actions:

@@

In total, we obtained 2354 distinct actions which summed up to a total
annotation duration of 145 minutes. On average, one child accounted for 92
actions (SD=23).

\begin{figure*}[ht!] 
    \centering 
    \includegraphics[width=0.9\linewidth]{domino-time-active.pdf} 
    \caption{\small \textbf{Number and type of actions for each run} (n=480,
        spontaneous actions, excluding \emph{explore}). Generally, the number of
        actions does not decrease over time (from run 1.1 to run 2.9).  The
        first 7 runs correspond to the \textit{expected phase}, the second 7
        runs correspond to the \textit{unexpected phase}. Especially during run
        2.3, the first time when the robot showed an unexpected behaviour,
        children tended to \textit{look} more at the experimenter. During the
        unexpected phase, also \textit{talk} and \textit{gesture} seem to be
        increased.}

    \label{fig:domino-time-active} 
\end{figure*}

The actions \textit{put, remove} and \textit{call} were ``requested'' actions
because the scenario required them and children were asked to carry them out.
Hence, these actions are not relevant to be analyzed quantitatively. The other
actions \textit{explore, misuse, gesture, touch, show, talk} and \textit{look}
are spontaneous actions that were not requested but arose in the interaction.
These actions reflect engagement, and are the one we considered during the
analysis. Figure~\ref{fig:domino-time-active} shows the distribution of these
actions over the different runs, in any condition.

@@

One pre-interview and two interviews (between run \emph{1.5} and \emph{2.1} and
at the end) were conducted with the children.  We set up the interviews like a
casual conversation / discussion, so we did not separate the two children in
order to keep the situation natural. We used casual and informal language
adapted to the age of the children, and encouraged them to justify their answers
and tell us more details by asking, for example, \textit{How do you know ... ?}
or \textit{Tell me, what have you observed when ... ?}. We paid attention to not
``put words in children's mouth''. Consequently, though we re-phrased and
repeated some questions, we never forced children to give an answer, and
accepted when they said they would not know or when they did not respond at all.

\begin{table}[h]
\centering
\footnotesize
\begin{tabular}{p{0.9\linewidth}}
    \toprule
    \textbf{Construct \emph{expectations}} \\
    \midrule

@@

    %%
    \toprule
    \textbf{Ascribe cognitive connections} \\
    \midrule

@@

    \caption{\small Constructs and questions used during the semi-structured interviews
    with children. The Ranger robot toy box is abbreviated with ``R''. Questions
    related to the construct \emph{expectations} were asked during the
    pre-interview.}

@@

In designing our interview script and selecting relevant questions, we took
inspiration from previous work on child-robot interaction and children's
perception of robots
\cite{kahn_jr._robotic_2006,weiss_i_2009,leite_influence_2013}. For instance, we
applied and adapted some of the \emph{constructs} and example questions from the
questionnaires used in \cite{kahn_jr._robotic_2006} and \cite{weiss_i_2009}. A
\emph{construct} addresses a specific factor (topic) that can be measured by
several questions. For instance, the construct ``cognitive connections'' (using
Flavell's terminology~\cite{flavell1988development}) considers the robot's
ability to hear and to see (perceptual skills), as attributed by the children.
The construct ``moral standing'' and the respective question was taken from
\cite{kahn_jr._robotic_2006}.\footnote{According to
\cite{kahn_jr._robotic_2006}, \textit{moral} refers to considerations based
on an artifact's physical or psychological welfare, and virtue (whether the
artifact deserves care). An attribution of moral standing reflects, for
instance, that the robot engenders moral regard, is morally responsible,
blameworthy, has rights or deserves respect.}			Similarly, we grouped
questions according to specific constructs that they evaluate (see
Table~\ref{tab:domino-questions}).

@@

The interviews were analyzed in a qualitative manner (we qualitatively
transcribed interviews, \ie we did not craft a full word-by-word transcript but
noted down any statement that was useful and relevant) and used to compute the
anthropomorphism index presented below.

@@

There are different possibilities to measure engagement in HRI, depending on the
specific research question and context. Metrics to measure behavioural engagement
in the interaction include, for instance, conversation analysis (\eg used in
\cite{short_no_2010}) or general attention analysis. These can be studied by
analyzing interaction videos in terms of head movement, eye tracking and gesture
/ body movement analysis (\eg in \cite{sidner_explorations_2005}).
Post-measurements to measure emotional engagement can be questionnaires that try
to assess constructs like the perceived presence and involvement in the
interaction. An example is the \textit{Interactive Experience Questionnaire}
(originally developed by \cite{lombard_measuring_2000}) of which adapted
versions were used in
\cite{kidd_effect_2004,bainbridge_effect_2008,short_no_2010}.

A mix of several methods has been used in a long-term interaction study with 8-9
year old children, \cite{leite_long-term_2013}. The author measured engagement
through video observations (by analyzing the amount of time that children spent
looking at the robot), interviews, and questionnaires. The interviews were
semi-structured, containing initial yes-or-no questions followed by open-ended
questions that allowed children to justify and elaborate their answers. We have
chosen a similar approach that considers both children's behavioural and
emotional engagement.

With 4-5 year old children, however, we cannot use rating scales to ask them
about constructs as abstract as social presence or how much they felt involved
in the interaction. Therefore we have attempted to combine \emph{interaction}
data with the results of the semi-structured interviews.  Several of the
aforementioned coded actions can reflect engagement: \textit{explore, misuse,
gesture, touch, show,} and \textit{talk}. Also, \textit{look} at the
experimenter can be considered as engagement in the interaction: We assume that
children look at the experimenter when they are surprised and seek for help.
This behaviour reflects that they want to make sense of the robot's behaviour and
that they notice it as ``strange''. Further, we can take into account how they
refer to Ranger and describe the robot (as well as their experience) in the
interviews.  Specifically, we used the constructs \textit{ascribe intention,
    ascribe cognitive abilities, ascribe emotional state}, and \textit{ascribe
    moral standing} (Table~\ref{tab:domino-questions}) to assess how far they
    anthropomorphize the robot.

To account for the fact that anthropomorphism arises in an interaction, we try
to bring both children's perception of the robot (post-measurement) and their
behaviour toward it (in-the-moment measurement) together. By doing so, we would
like to obtain a qualitative \textbf{anthropomorphism
index}~\cite{fink2014dynamics}. We quantify the index by attributing points for each
anthropomorphic \emph{perception} of the robot (max.  13 points) and for specific kinds
of human-like \emph{behaviour} toward the robot (max.  3 points). The index included the
following aspects:\\

\textbf{Perception: \textit{(max. 13 points)}}
\begin{itemize}
	\item Ascribe \textbf{mental states / feelings} to Ranger: \textit{0-4 points}\\
	(2 points for agreeing that Ranger can be happy or sad; 2 points for attributing Ranger with hunger or tiredness)
	\item Ascribe \textbf{cognitive abilities / intention} to Ranger: \textit{0-4 points}\\ 
	(each 0.5 points for ascribing seeing and hearing ability; 1 point for agreeing that Ranger can go out the door by itself; 1 point for disagreeing that Ranger always obeys; 1 point for agreeing that Ranger can do something silly)
	\item Ascribe \textbf{sociality / companionship} to Ranger: \textit{1 point}\\ 
	(1 point for agreeing that Ranger can be a friend)
	\item Ascribe \textbf{moral standing} to Ranger: \textit{1 point}\\ (1 point for disagreeing that Ranger be left alone at home)
	\item Other \textbf{anthropomorphic statements}: \textit{0-3 points}\\ 
	(1 point for anthropomorphic reason for Ranger's misbehaviour; 2 points for anthropomorphic reason for not leaving Ranger alone) 
\end{itemize}

\textbf{behaviour: \textit{(max. 3 points)}}
\begin{itemize}
	\item Use of \textbf{direct speech} toward Ranger: \textit{1 point}\\
	(\textit{calling} the robot to come over is not considered)
	\item Use of \textbf{polite formulations} toward Ranger: \textit{1 point}\\ 
	(\eg saying \textit{``thank you Ranger''} or \textit{``please Ranger ...''} or \textit{``goodbye''})
	\item Use of \textbf{social or pointing gestures} toward Ranger: \textit{1 point}\\ 
	(\eg waving at the robot, nodding)
\end{itemize}
